Compositional equivalences based on Open pNets\thanks{This work was partially funded by the Associated Team FM4CPS between INRIA and ECNU, Shanghai}

### 2020-01-03 2016-Weak-FH-Bisim/WeakBisim.tex
nettoyages divers

\TODO{Explain the partial OA}
\end{example}

@@

\subsection{Weak FH-bisimulation}
For defining a bisimulaiton relation between weak open automatas, two options are possible. Either we define a simulation similar to the strong simulation but based on open automata, this would look like the FH-simulation but would need to be adapted to weak open transitions. Or we define directly and classically a weak FH-simulation as a relation between two open automata, relating the open transition of one open automaton with the transition of the weak open automaton derived from the second one.

@@

Two pNets are weak FH-bisimilar if there exist a relation between their associated 
automata that is a Weak FH-bisimulation and their initial states are in the relation, i.e. 
the predicate associated to the relation between the initial states is \True.
 \end{definition}

@@

    \bigskip
    \ERIC{Anciennes versions:}
$\ostate{1000}\OTarrow{\underline{send(m,b)}, \phi_1 = [x=s_1\_b \land
    m=s_1\_m]}\ostate{2100}$\\
and $\qquad\ostate{2100}$ variable assignments are: $\psi_1 = \{m_1\_m:=m; m_1\_x:=x; s_2\_b:=s_1\_b\}$;\\
\\
$\ostate{2100}\OTarrow{\tau <m\_send(m_1\_m,m_1\_x), r\_rec(?m',?b)>,
  \phi_2 = [r_0\_b=b \land m'=m_1\_m \land b=m_1\_x]}\ostate{2001}$\\
and $\qquad\ostate{2001}$ assignments: $\psi_2 = \{r_1\_b:=r_0\_b; r_1\_m:=m;
; s_2\_b:=s_1\_b\}$;\\
\\
$\ostate{2001}\OTarrow{v,
  \phi_3 = [v=out(r_1\_m)]}\ostate{2002}$\\
and $\qquad\ostate{2002}$ variablesassignments: $\psi_3 = \{r_2\_b:=r_1\_b; s_2\_b:=s_1\_b\}$;\\

@@

\ERIC{
=> Context VCE, ref to FASE paper\\
=> Current prototype: programatic creation of pNet objects, through the pNet API in VCE. On the middle term, this should be compiled from some language or graphical formalism.\\
=> Open transitions are computed as a direct implementation of the algorithm above. Steps 2a-2b are merged: at each pNet node, after application of rule \TrDeux, the resulting OT is built from the premisses, without explicitely constructing the deduction tree. \\
=> Variable names are generated for all ``fresh'' and ``clone'' operations, using structured variable names ensuring uniqueness of fresh names, but also reasonable readability of these names for debugging purposes.\\
=> Satisfiability check is done only at toplevel of the deduction tree construction. An alternative would be to submit the satisfiability check to the SMT solver at each level of the tree construction, potentialy reducing the overall number of combinations. But submission to the SMT engine is costly, and more complexity analysis is required before deciding if this would be worthwhile.\\
=> Simplification is mostly implemented. It is not strictly required for the Open Automaton construction, but it will be critical later for predicate comparison in the bisimulation algorithm.}

Independently, it is clear that most interesting properties of
such complex systems will not be provable by strong bisimulation.
Next steps will include the investigation of weak
versions of the FH-bisimulation, using the notion of
\emph{synchronised actions} mentionned in the paper.
%We are convinced that this work will contribute to the design of tools for
%verifying open distributed systems.

@@

\begin{figure}[t]
   \includegraphics[width=5.5cm]{XFIG/SimpleProt-Schema}
   \includegraphics[width=6.5cm]{XFIG/SimpleProt2-Spec}
   \caption{Specification pNet Schema and Specification}
   \label{App:SimpleProt:Spec}

\begin{figure}[t]
  \centerline{\includegraphics[width=12cm]{XFIG/SimpleProt2-pNet}}
  \caption{Composed pNet with the Simple Protocol Implementation}  \label{App:SimpleProt:Impl}
\end{figure}

The first piece of code is the textual definition of the SimpleProtSec
pNet. This code should be intuitive enough to read, with the following
conventions:

\begin{itemize}
  \item Constants of any type (including Action) must be declared as
    ``const''. They are used either as functions with argument, as
    typically \texttt{in(msg)}, or constants without argument, denoted
    as \texttt{\$tau}. Identifiers without the ``?'' marker are
    variables.
  \item Variables can be declared either: inside a pLTS state (normal
    state variable), as global variable of a pLTS (see e.g. ...), or
    as an input variable in a transition, as \texttt{?msg} in pLTS M1
    here.
  \item \TODO{continue...}
\end{itemize}

\TODO{keep ?}
\paragraph{Variable management.}
The variables in each synchronisation vector are considered local:
for a given pNet expression, we must have fresh local variables for
each occurrence of a vector (i.e. each time we instantiate rule
\TrDeux). Similarly the state variables of each copy of a
given pLTS in the system, must be distinct, and those created for each
application of \TrDeux\ have to be fresh and all distinct. 
This is implemented within the open-automaton generation algorithm,
using name generation using a global counter as a suffix.

@@

From this code,A more compact view (maybe more readable ???) is given
by the pretty-printer of VerCors:

\begin{lstlisting}[basicstyle=\scriptsize\ttfamily, language=java, frame=single]
SimpleProtImpl2
--------------------
SV6 SVs: <p_send(m), in(m), _>-->in(m)
SV7 SVs: <p_a, _, _>-->p_a, p_a!=(p_send(m))
SV8 SVs: <_, out(m, c), q_recv(m, c)>-->out(m, c)
SV9 SVs: <_, _, q_a>-->q_a, q_a!=(q_recv(m, c))
SV11 SVs: <_, _tau_, _>-->_tau_
Holes: P
Subnets: medium
- medium
- SV0 SVs: <in(m), _, _>-->in(m)
- SV1 SVs: <s_send(m, c), m_recv(m, c), _>-->_tau_
- SV2 SVs: <_, m_send(m), r_recv(m)>-->_tau_
- SV3 SVs: <s_ack, _, r_ack>-->_tau_
- SV4 SVs: <s_error, m_error, _>-->_tau_
- SV5 SVs: <_, _, out(m, c)>-->out(m, c)
- SV10 SVs: <_, _tau_, _>-->_tau_
- pLTS: Sender
- - Transition: s0---in(msg)-->s1, { s1_msg := msg s1_c := 0}
- - Transition: s1---s_send(s1_msg, s1_c)-->s2, {}
- - Transition: s2---s_ack-->s0, {}
- - Transition: s2---s_error-->s1, { s1_c := s2_c+1}
- - Variable: s1_c, State: s1, Value(s) = 0,s2_c+1,
- - Variable: s1_msg, State: s1, Value(s) = msg,
- pLTS: M1
- - Transition: a0---m_recv(msg, c)-->a1, { a1_msg := msg a1_c := c}
- - Transition: a1---m_send(a1_msg, a1_c)-->a0, {}
- - Transition: a1---_tau_-->a2, {}
- - Transition: a2---m_error-->a0, {}
- - Variable: a1_msg, State: a1, Value(s) = msg,
- - Variable: a1_c, State: a1, Value(s) = c,
- pLTS: Receiver
- - Transition: r0---r_recv(msg, c)-->r1, { r1_msg := msg r1_c := c}
- - Transition: r1---out(r1_msg, r1_c)-->r2, {}
- - Transition: r2---r_ack-->r0, {}
- - Variable: r1_msg, State: r1, Value(s) = msg,
- - Variable: r1_c, State: r1, Value(s) = c,
Holes: Q
\end{lstlisting}

>>>>

\TODO{Eric: Explain the partial OA}
\end{example}

@@

\subsection{Weak FH-bisimulation}
For defining a bisimulation relation between weak open automata, two options are possible. Either we define a simulation similar to the strong simulation but based on open automata, this would look like the FH-simulation but would need to be adapted to weak open transitions. Or we define directly and classically a weak FH-simulation as a relation between two open automata, relating the open transition of one open automaton with the transition of the weak open automaton derived from the second one.

@@

Two pNets are weak FH-bisimilar if there exists a relation between their associated 
automata that is a Weak FH-bisimulation and their initial states are in the relation, i.e. 
the predicate associated to the relation between the initial states is \True.
 \end{definition}

@@

    \bigskip
    ...
    \bigskip

@@

\ERIC{
  => Context VCE, ref to FASE paper, citer Avocs et  PEPM\\
=> Simplification is mostly implemented. It is not strictly required for the Open Automaton construction, but it will be critical later for predicate comparison in the bisimulation algorithm.}

@@

%% \begin{figure}[t]
%%    \includegraphics[width=5.5cm]{XFIG/SimpleProt-Schema}
%%    \includegraphics[width=6.5cm]{XFIG/SimpleProt2-Spec}
%%    \caption{Specification pNet Schema and Specification}
%%    \label{App:SimpleProt:Spec}

%% \end{figure}

%% \begin{figure}[t]
%%   \centerline{\includegraphics[width=12cm]{XFIG/SimpleProt2-pNet}}
%%   \caption{Composed pNet with the Simple Protocol Implementation}  \label{App:SimpleProt:Impl}
%% \end{figure}

The first piece of code is the textual definition of the SimpleProtSec
pNet, that was drwan in Figure \ref{SimpleProt:Impl}, page \pageref{SimpleProt:Impl}. This code should be intuitive enough to read, with the following
conventions, that brings some user-friendly features that are maps by the editor into pure pNet constructs.

\begin{itemize}
  \item Constants of any type (including Action) must be declared as
    ``const''. They are used either as functions with argument, as
    typically \texttt{in(msg)}, or constants without argument, denoted
    as \texttt{\$tau}. Identifiers without the ``?'' marker are
    variables.
  \item Variables can be declared either: inside a pLTS state (normal
    state variable), as global variable of a pLTS (see e.g. ...), or
    as an input variable in a transition, as \texttt{?msg} in pLTS M1
    here.
  \item the variables in the guards of synchronisation vectors do not need to be explicitely quantified: all variable in a guard that does not appear inside the vector actions will be recognised as bound by a \emph{foral} quantifier inside the guard.
    \item the tools will check that everything is correctly declared, and that variables are used properly, that vectors have coherent length, etc.
\end{itemize}

%% \TODO{keep ?}
%% \paragraph{Variable management.}
%% The variables in each synchronisation vector are considered local:
%% for a given pNet expression, we must have fresh local variables for
%% each occurrence of a vector (i.e. each time we instantiate rule
%% \TrDeux). Similarly the state variables of each copy of a
%% given pLTS in the system, must be distinct, and those created for each
%% application of \TrDeux\ have to be fresh and all distinct. 
%% This is implemented within the open-automaton generation algorithm,
%% using name generation using a global counter as a suffix.

@@

%% From this code,A more compact view (maybe more readable ???) is given
%% by the pretty-printer of VerCors:

%% \begin{lstlisting}[basicstyle=\scriptsize\ttfamily, language=java, frame=single]
%% SimpleProtImpl2
%% --------------------
%% SV6 SVs: <p_send(m), in(m), _>-->in(m)
%% SV7 SVs: <p_a, _, _>-->p_a, p_a!=(p_send(m))
%% SV8 SVs: <_, out(m, c), q_recv(m, c)>-->out(m, c)
%% SV9 SVs: <_, _, q_a>-->q_a, q_a!=(q_recv(m, c))
%% SV11 SVs: <_, _tau_, _>-->_tau_
%% Holes: P
%% Subnets: medium
%% - medium
%% - SV0 SVs: <in(m), _, _>-->in(m)
%% - SV1 SVs: <s_send(m, c), m_recv(m, c), _>-->_tau_
%% - SV2 SVs: <_, m_send(m), r_recv(m)>-->_tau_
%% - SV3 SVs: <s_ack, _, r_ack>-->_tau_
%% - SV4 SVs: <s_error, m_error, _>-->_tau_
%% - SV5 SVs: <_, _, out(m, c)>-->out(m, c)
%% - SV10 SVs: <_, _tau_, _>-->_tau_
%% - pLTS: Sender
%% - - Transition: s0---in(msg)-->s1, { s1_msg := msg s1_c := 0}
%% - - Transition: s1---s_send(s1_msg, s1_c)-->s2, {}
%% - - Transition: s2---s_ack-->s0, {}
%% - - Transition: s2---s_error-->s1, { s1_c := s2_c+1}
%% - - Variable: s1_c, State: s1, Value(s) = 0,s2_c+1,
%% - - Variable: s1_msg, State: s1, Value(s) = msg,
%% - pLTS: M1
%% - - Transition: a0---m_recv(msg, c)-->a1, { a1_msg := msg a1_c := c}
%% - - Transition: a1---m_send(a1_msg, a1_c)-->a0, {}
%% - - Transition: a1---_tau_-->a2, {}
%% - - Transition: a2---m_error-->a0, {}
%% - - Variable: a1_msg, State: a1, Value(s) = msg,
%% - - Variable: a1_c, State: a1, Value(s) = c,
%% - pLTS: Receiver
%% - - Transition: r0---r_recv(msg, c)-->r1, { r1_msg := msg r1_c := c}
%% - - Transition: r1---out(r1_msg, r1_c)-->r2, {}
%% - - Transition: r2---r_ack-->r0, {}
%% - - Variable: r1_msg, State: r1, Value(s) = msg,
%% - - Variable: r1_c, State: r1, Value(s) = c,
%% Holes: Q
%% \end{lstlisting}
